---
title: "Lab 09: Annotating Data Visualizations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
theme_set(theme_minimal())
```

## Pick a dataset

Today you'll have a choice of three datasets to work with. You can choose
either to investigate Donald Trump's approval rating:

```{r}
trump <- read_csv("https://statsmaths.github.io/stat_data/trump_approval.csv")
```

Murder rates across different cities in the US:

```{r}
murder <- read_csv("https://statsmaths.github.io/stat_data/murder_by_city.csv")
```

Or the daily close values of the Dow Jones Industrial Average:

```{r}
djia <- read_csv("https://statsmaths.github.io/stat_data/djia_daily.csv")
```

Pick a dataset (feel free to delete the others from the Rmd file if you would
like), and work on the instructions in the next section.

## Instructions

Your goal in this lab is to construct an interesting visualization with the
data that you choose above. You should make use of several manual annotations
in an effort to tell a visual story about the dataset. Below, you should
eventually provide a few sentences describing what you found in the dataset
and then a single plot illustrating the interesting aspects of the data.
You should incorporate some external information into your plot.

Note that you will certainly want to create many initial exploratory plots
before settling on a final one for analysis. This time, make sure you that you
upload both the Rmd and the html files to GitHub (remember, the final report
should have just one single visualization).

Also note that you will have to use the format `x = as.Date("2017-01-06")` to
add annotations to an axis that uses time as a component.

## Analysis
The plot below shows a representation of the murders in various cities in 2014 compared to the murders in those same cities in 2015. The diference in murders is represented by the color gradient, with yellow showing an increase in murders and blue showing a decrease. For the most part, cities with high murder rates in 2014 seemed to fail to lower their rates in 2015, while cities with low murder rates more effectivley lowered murder rates. Specififc points to notice are New York and Chicago, which had particuarly high murder rates compared to the rest of the country, Baltimore, which had significantly higher increase in murders rates as compared to the rest of the country. Also worth noting is the addition of Richmond's murder data, added in red.
```{r}
library(viridis)
 ggplot(murder, aes(murders_2014, murders_2015)) +
  geom_point(aes(color = change)) +
  scale_color_viridis() +
  ylab("Murders in 2015") +
  xlab("Murders in 2014") +
  ggtitle("Change in Murders Between 2014 and 2015") +
  annotate("text", x = 210, y = 360, label = "Baltimore") +
  annotate("text", x = 330, y = 365, label = "New York") +
  annotate("text", x=417, y = 490, label = "Chicago") +
  annotate("point", x = 41, y = 42, color = "red") +
  annotate("text", x = 62, y = 30, label = "Red = Richmond", size = 3)
```









