---
title: "Premiee League Goals"
author: "Sam Zielinski"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
library(smodels)
```

## Introduction
The data that I collected is from the Premier League, in England. I enjoy watching soccer, and so though that it would be cool to look at the effect that home and away games has on the amount of goals scored and the amount of shots taken. Simaraly I wanted to look at the trend between shots taken and goals scored. Having the home field in a soccer game, or any sport for that matter, is generally seen as a big advantage to the home team, do to having more fans and a familiarity to the field. I believe that this advantage will cause the home team to both score more goals, as well as have more shots per game. I also predict that more shots per game means more goals for that team. 

## Methods
To collect the data I went to WhoScored.com, a website that collects a variety of data and statistics from games played in the top soccer leagues all over the world. I looked at the match summury for every Premier League game played in the past seven weeks, and took the amount of goals scored by each team per game, whether the team was home or away, and how many shots the team had. I then created a three linear regression models, one showing the average goal with respect to being home or away, one showing average shots with respect to being home or away, and one showing goals scored as a function of shots taken. I then ran each model through a regression table with a 95% confidence level. 

## Results
```{r}
goals <- read_csv("https://raw.githubusercontent.com/urstats/stat209-s18-samueltz/master/goals.csv")
model <- lm_basic(goals_scored ~ 1 + home_away, data = goals)
reg_table(model, level = 0.95)

goals_home_away <- group_summarize(goals, home_away)

ggplot(goals_home_away, aes(home_away, goals_scored_mean)) +
  geom_col(fill = "steelblue4") +
  xlab("Home or Away") +
  ylab("Average Goals Scored")
  
```
This model shows the effect that being home or away had on the average number of goals scored per team per game. According to the regression there is no significant realtionship between the two variables, becuase the model redicts that the mean goals scored for home and away games may overlap within a 95% confidence level.
```{r}
model <- lm_basic(shots ~ 1 + home_away, data = goals)
reg_table(model, level = 0.95)

ggplot(goals_home_away, aes(home_away, shots_mean)) +
  geom_col(fill = "steelblue4") +
  xlab("Home or Away") +
  ylab("Average Shots per Game")
```
The second model shows the effect of being home or away on the amount of shots each team had per game. According to the regression, and with a 95% confidence level, teams at home have on average 1.295  to 5.334 more shots per game more than teams that are playing away. 

```{r}
model <- lm_basic(goals_scored ~ 1 + shots, data = goals)
reg_table(model, level = 0.95)

ggplot(goals, aes(shots, goals_scored)) +
  geom_point() +
  geom_smooth(method = "lm")
```
The final model show hoe the amount of shots taken effect the amount of goals scored by each team. The regression shows that, with 95% confidence, for each additional shot taken, the amount of goals scored increases by about 0.06391. 


## Conclusions
The models showed that there is a significant direct correlation between the amount of shots taken and goals scored by each team in a given game, as well as a significant and direct correlation between the amount of shots taken and whether the team was home or not. There was no signifigance as to whether being home or awa influenced the amount of goals scored. Overall this means that the teams that play at homehave more shots on average than teams that play away, however teams that play at home don't necessarily score more goals than teams that are away. It was also confirmed that the more shots a team took, the more goals they normally scored. 


## References
www.whoscored.com



